### 2017-02-07 2017mergingmerges/section/XX-Processes.tex
hide iota combinator and abstract group

% BL: Just start into the description.
%     Readers will work out what's coming by flipping through the paper themselves.
% This section describes processes in detail. We start with an informal explanation of processes before formally defining the syntax and dynamic semantics.

A process defines a stream computation, taking any number of input streams and producing at least one output stream.
Before describing the @group@ process, we start by looking at one of the simplest combinators, @iota@, which produces a stream of increasing numbers.
It takes no inputs, and produces one output stream @xs@.
Each process has its own local heap where the values are stored, and in this case we initialise the local variable @i@ to @0@.
This variable will be incremented and pushed.
Each process also has a current label, which denotes the instruction to perform next.
The initial label for @iota@ is @L0@.
Each process has a mapping from labels to instructions.
In this case we have two instructions, @L0@ and @L1@.
The instruction for @L0@ pushes the current value of variable @i@ onto the output stream, then proceeds to move to label @L1@.
Instruction @L1@ moves back to @L0@, while also incrementing the variable @i@.

\begin{code}
process (iota)
     ins: 
    outs: xs
    heap: {i = 0}
   label: L0
  blocks: L0 = push xs i  L1
          L1 = jump       L0{i = i + 1}
\end{code}

When executed, this program produces an infinite stream of increasing numbers: $0, 1, 2\ldots$ while the label alternates between @L0@ and @L1@.

The @uniques = group file1@ is a more interesting example.
Recall that @group@ removes consecutive duplicates from its input stream.
It has one input, @file1@, and one output, @uniques@.
There are three variables in the heap: @first@ is initialised to @True@ as the first pulled value has no last value to compare against; @value@ stores the most recently pulled value, and @last@ stores the last pulled value.
The @last@ and @value@ variables are initialised to $0$, but could be initialised to any value: these initial values will not be used.
The initial label is @L0@, which pulls from the input @file1@. This blocks waiting for an input value, and when one is received, it is stored in @value@ and the process moves to @L1@.
Instruction @L1@ performs a case analysis on a boolean: if it is the first read value, or the last value is not equal to the most recent value, it jumps to @L2@; otherwise it jumps to @L3@.
Instruction @L2@ pushes the most recent value to the output and jumps to @L3@, updating the last value to the most recent value, and setting first to @False@.
Finally, the instruction for @L3@ `drops' the recently pulled value from @file1@ and jumps back to @L0@.
This dropping is required to coordinate between multiple processes that read from the same input: now that the read value from @file1@ has been dropped, another process is free to pull the next value from @file1@ if it so wishes.

\begin{code}
process (group)
     ins: file1
    outs: uniques
    heap: {first = True, last = 0, value = 0}
   label: L0
  blocks: L0 = pull file1   value               L1
          L1 = case (or first (last /= value))  L2               L3
          L2 = push uniques value               L3{last = value, first = False}
          L3 = drop file1                       L0
\end{code}

@@

>>>>

Each combinator is defined as a stream process and expressed in terms of a simple imperative program with a local heap. The process pulls data from an aribrary number if input streams and pushes data to at least one output stream. The definition of the @group@ combinator which removes consecutive elements from its input stream is as follows:

\begin{alltt}
 group 
   = \(\lambda\) (s1: Stream a) (s2: Stream a). 
     \(\exists\) (first: Bool)  (last: a) (value: a).
\end{alltt}\vspace{-1em}
\begin{code}
     process
     { ins:    { s1 }
     , outs:   { s2 }
     , heap:   { first = True, last = 0, value = 0 }
     , label:  L0
     , instrs: { L0 = pull s1 value                    L1 {}
               , L1 = case (first || (last /= value))  L2 {}  L3 {}
               , L2 = push s2 value                    L3 { last = value, first = False }
               , L3 = drop s1                          L0 {} } }
\end{code}

The @group@ combinator has two parameters, @s1@ and @s2@ which are the bind the input and output streams respectively. The combinator also defines three local variables @first@, @last@ and @value@. The body of the combinator is a record that defines the process. The @ins@ field of the record defines the set of input streams and the @outs@ field the set of output streams. The @heap@ field gives the starting state of each of the local variables. The @instrs@ field contains a set of labeled instructions that define the program, while the @label@ field gives the label of the first instruction.

Recall that @group@ removes consecutive duplicates from its input stream.
It has one input, @file1@, and one output, @uniques@.
There are three variables in the heap: @first@ is initialised to @True@ as the first pulled value has no last value to compare against; @value@ stores the most recently pulled value, and @last@ stores the last pulled value.
The @last@ and @value@ variables are initialised to $0$, but could be initialised to any value: these initial values will not be used.
The initial label is @L0@, which pulls from the input @file1@. This blocks waiting for an input value, and when one is received, it is stored in @value@ and the process moves to @L1@.
Instruction @L1@ performs a case analysis on a boolean: if it is the first read value, or the last value is not equal to the most recent value, it jumps to @L2@; otherwise it jumps to @L3@.
Instruction @L2@ pushes the most recent value to the output and jumps to @L3@, updating the last value to the most recent value, and setting first to @False@.
Finally, the instruction for @L3@ `drops' the recently pulled value from @file1@ and jumps back to @L0@.
This dropping is required to coordinate between multiple processes that read from the same input: now that the read value from @file1@ has been dropped, another process is free to pull the next value from @file1@ if it so wishes.

@@

% BL: I don't think describing iota works at this point. This combinator is not used in the motivating example, so skipping to it seems disjointed.

% Before describing the @group@ process, we start by looking at one of the simplest combinators, @iota@, which produces a stream of increasing numbers.
% It takes no inputs, and produces one output stream @xs@.
% Each process has its own local heap where the values are stored, and in this case we initialise the local variable @i@ to @0@.
% This variable will be incremented and pushed.
% Each process also has a current label, which denotes the instruction to perform next.
% The initial label for @iota@ is @L0@.
% Each process has a mapping from labels to instructions.
% In this case we have two instructions, @L0@ and @L1@.
% The instruction for @L0@ pushes the current value of variable @i@ onto the output stream, then proceeds to move %to label @L1@. Instruction @L1@ moves back to @L0@, while also incrementing the variable @i@.

% \begin{code}
% process (iota)
%      ins: 
%     outs: xs
%     heap: {i = 0}
%    label: L0
%   blocks: L0 = push xs i  L1
%           L1 = jump       L0{i = i + 1}
% \end{code}

% When executed, this program produces an infinite stream of increasing numbers: $0, 1, 2\ldots$ while the label alternates between @L0@ and @L1@.

% The @uniques = group file1@ is a more interesting example.
